### 2022-03-05 data-raw/00-Toy-Example-Data-Preparation.Rmd

---
title: "Toy example dataset"
output: 
---
In this notebook we provide a toy dataset that is used to calculate both conventional accessibility and _spatial available_ accessibility. This is a more smaller and theoretical alternative dataset that can be used in place the the real-world TTS-16.

Libraries:
```{r message=FALSE}
library(sf)
library(tidyverse)
```

Load data for example:
```{r}
load("toy-example-data-inputs/od_table.rda")
toy_od_table <- od_table
load("toy-example-data-inputs/simulated_data.rda")
toy_sim_zones <- simulated_data
load("toy-example-data-inputs/trips.rda")
toy_trips <- trips
```

Remove unneeded columns and add trips to OD:
```{r}
toy_od_table <- toy_od_table %>% 
  dplyr::select(-c("f", "V_ij", "W_ij", "V_ij_r")) %>%
  merge(toy_trips %>%
          dplyr::select(Origin, Destination, trips),
               by = c("Origin", "Destination"), all.x=TRUE)
```

Save it to /data folder:
```{r}
usethis::use_data(toy_od_table, overwrite = TRUE)
usethis::use_data(toy_sim_zones, overwrite = TRUE)
```