### 2022-03-11 data-raw/00-TTS16-and-Bound-Data-Preparation-OD.Rmd

---
title: "TTS Origin-Destination Data by Full-Time Employment"
output: 
---

In this notebook we process a cross-tabulation of origin and destinations by people employed full-time in the Greater Golden Horseshoe (GGH) in 2016.

The source data is in file `tts-2016-OD-FT-Employment.txt`, which was retrieved from the Transportation Tomorrow Survey Data Retrieval System on October 28, 2021.

The centroids of all work zones (origins) and full-time (FT) employment zones (destinations) are calculated.

## Preliminaries

Load packages:
```{r load-packages}
library(readxl)
library(sf)
library(tidyverse)
library(units)
```

## Traffic Analysis Zones

Read the boundaries of the traffic analysis zones:
```{r}
ggh_taz <- st_read("TTS16-data-inputs/OD-by-FT-Employment/tts06_83_region.shp")
```

Project object:
```{r}
ggh_taz <- ggh_taz %>%
  st_transform(crs = 32617)
```

Prepare traffic analysis zones:
```{r}
ggh_taz <- ggh_taz %>%
  transmute(GTA06 = as.character(GTA06),
            AREA = st_area(geometry) %>% 
              set_units(km^2) %>% 
              drop_units())
```

Extract the IDs for all TAZ:
```{r}
ghta_taz_id <- ggh_taz$GTA06
```

## Planning boundaries

Read the boundaries of the traffic analysis zones:
```{r}
ggh_pd <- st_read("TTS16-data-inputs/Boundaries/tts06_pd_83.shp")
```

Project object:
```{r}
ggh_pd <- ggh_pd %>%
  st_transform(crs = 32617)
```

Adding region names to the object as they are not included in the origin data:
(http://dmg.utoronto.ca/pdf/tts/2016/2016TTS_Conduct.pdf):
```{r}
ggh_pd <- ggh_pd %>%
  mutate(REGION_name = 
           ifelse(REGION == 1, "Toronto", 
                  ifelse(REGION == 2, "Durham", 
                         ifelse(REGION == 3, "York", 
                                ifelse(REGION == 4, "Peel", 
                                       ifelse(REGION == 5, "Halton", 
                                              ifelse(REGION == 6, "Hamilton", 
                                                     ifelse(REGION == 11, "Niagara", 
                                                            ifelse(REGION == 12, "Waterloo", 
                                                                   ifelse(REGION == 13, "Guelph", 
                                                                          ifelse(REGION == 14, "Wellington", 
                                                                                 ifelse(REGION == 15, "Orangeville", 
                                                                                        ifelse(REGION == 16, "Barrie", 
                                                                                               ifelse(REGION == 17, "Simcoe", 
                                                                                                      ifelse(REGION == 18, "Kawartha Lakes", 
                                                                                                             ifelse(REGION == 19, "Peterborough City", 
                                                                                                                    ifelse(REGION == 20, "Peterborough County", 
                                                                                                                           ifelse(REGION == 21, "Orillia", 
                                                                                                                                  ifelse(REGION == 22, "Dufferin", 
                                                                                                                                         ifelse(REGION == 23, "Brantford", 
                                                                                                                                                ifelse(REGION == 24, "Brant", "NOOOO")))))))))))))))))))))
```

## Origin destination matrices for work and shop

Read trip tables. The query was for "trips" by zone of residence, zone of work, and trip purpose (Work):
```{r}
od_ft <- read_delim(file = "TTS16-data-inputs/OD-by-FT-Employment/tts-2016-OD-FT-Employment.txt", 
                      delim = "\t", 
                      col_names = FALSE)
```

Find the position in the table where the information on work trips begins:
```{r}
idx <- which(od_ft$X1 == "TABLE    : emp_stat (Full time)")
```

Slice the table to obtain the work trips (remove heading information):
```{r}
od_ft <- od_ft %>% 
  slice((idx+2):n())
```

Separate the zone identifiers and the trips, and convert to numeric:
```{r}
od_ft <- od_ft %>%
  separate(X1, into = c("Zones", "Persons"), sep = " (?=[^ ]+$)") %>%
  mutate(Zones = str_trim(Zones, side = "both")) %>%
  separate(Zones, into = c("Origin", "Destination"), sep = " (?=[^ ]+$)") %>%
  mutate(Origin = str_trim(Origin),
         Destination = str_trim(Destination),
         Persons = as.numeric(Persons))
```

The resulting object `od_ft` has the number of individuals who are employed full-time by place of residence (Origin) and place of work (Destination) using the GTA06 zoning system.

## Workers and Jobs

The od table can be used to calculate the workers and jobs by traffic analysis zone:
```{r}
workers_ft <- od_ft %>%
  group_by(Origin) %>%
  summarize(workers = sum(Persons),
            .groups = "drop") %>%
  transmute(GTA06 = Origin,
            workers)

jobs_ft <- od_ft %>%
  group_by(Destination) %>%
  summarize(jobs = sum(Persons),
            .groups = "drop") %>%
  transmute(GTA06 = Destination,
            jobs)

```

Join workers and jobs to simple features object:
```{r}
ggh_taz <- ggh_taz %>%
  # Join workers
  left_join(workers_ft,
            by = "GTA06") %>%
  # Join jobs
  left_join(jobs_ft,
            by = "GTA06") %>%
  # Replace NAs with zeros
  mutate(workers = replace_na(workers, 0),
         jobs = replace_na(jobs, 0))
```

Plot worker population in the ggh:
```{r}
ggplot() +
  geom_sf(data = ggh_taz,
          aes(fill = workers),
          color = NA) +
  scale_fill_fermenter(direction = 1)
```

Plot job distribution in the ggh:
```{r}
ggplot() +
  geom_sf(data = ggh_taz,
          aes(fill = jobs),
          color = NA) +
  scale_fill_fermenter(direction = 1)
```

## Creating input for travel-time OD calculations

Create centroids for each feature in ggh_taz:
```{r}
ggh_taz_cent <- st_centroid(ggh_taz) 
```

Attach the centroid point geometry to the origin object (workers) and the destination object (jobs)
```{r}
#workers (origins)
workers_ft_cent <- workers_ft %>% merge(ggh_taz_cent %>% dplyr::select(GTA06), by="GTA06") %>% st_sf()

work_origins <- cbind(workers_ft_cent, st_coordinates(st_transform(workers_ft_cent, crs = 4326))) %>%
  dplyr::rename(lon = "X", lat = "Y", id = "GTA06") %>% as.data.frame()  %>%
  dplyr::select(id, lon, lat) %>% as.data.frame()

```

```{r}
#jobs (destinations)
jobs_ft_cent <- jobs_ft %>% merge(ggh_taz_cent %>% dplyr::select(GTA06), by="GTA06") %>% st_sf

job_destinations <- cbind(jobs_ft_cent, st_coordinates(st_transform(jobs_ft_cent, crs = 4326))) %>%
  dplyr::rename(lon = "X", lat = "Y", id = "GTA06") %>% as.data.frame()  %>%
  dplyr::select(id, lon, lat)

```

## saving inputs for OD travel time calculations
```{r}
save("work_origins", file = "TTS16-data-inputs/Travel-Time-Calculations/inputs/work_origins.Rdata")
save("job_destinations", file = "TTS16-data-inputs/Travel-Time-Calculations/inputs/job_destinations.Rdata")
```

## saving inputs for later processing
```{r}
save("ggh_taz", file = "TTS16-data-inputs/OD-by-FT-Employment/ggh_taz.Rdata")
save("ggh_pd", file = "TTS16-data-inputs/Boundaries/ggh_pd.Rdata")
save("od_ft", file = "TTS16-data-inputs/OD-by-FT-Employment/od_ft.Rdata")
```