### 2022-10-24 Manuscript-Data-Package/Manuscript-Data-Package-cover-page.Rmd
submitted revision to journal. submission package in 'editorial-correspondance'.

---
title: "TTS2016R: A data set to study population and employment patterns from the 2016 Transportation Tomorrow Survey (TTS) in the Greater Golden Horseshoe Area, Ontario, Canada"
author:
  - name: "Anastasia Soukhov"
    email: "soukhoa@mcmaster.ca"
    affiliation: SEES
  - name: "Antonio Páez"
    email: "paezha@mcmaster.ca"
    affiliation: SEES
    footnote: 1
address:
  - code: SEES
    address: "School of Earth, Environment and Society, McMaster University, Hamilton, ON, L8S 4K1, Canada"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "This paper describes and visualises the data contained within the {TTS2016R} data package created in `R`, the statistical computing and graphics language. In addition to a synthetic example, {TTS2016R} contains home-to-work commute information for the Greater Golden Horseshoe (GGH) area in Canada retrieved from the 2016 Transportation Tomorrow Survey (TTS). Included are all Traffic Analysis Zones (TAZ), the number of people who are employed full-time per TAZ, the number of jobs per TAZ, the count of origin destination (OD) pairs and trips by mode per origin TAZ, calculated car travel time from TAZ OD centroid pairs, and associated spatial boundaries to link TAZ to the Canadian Census. To illustrate how this information can be analysed to understand patterns in commuting, we estimate a distance-decay curve (i.e., impedance function) for the region. {TTS2016R} is a growing open data product built on `R` infrastructure that allows for the immediate access of home-to-work commuting data alongside complimentary objects from different sources. The package will continue expanding with additions by the authors and the community at-large by requests in the future. {TTS2016R} can be freely explored and downloaded in the associated [Github repository](https://github.com/soukhova/TTS2016R) where the documentation and code involved in data creation, manipulation, and all open data products are detailed."

keywords: Jobs; population; work; commute; travel time; impedance; Greater Toronto and Hamilton Area; Greater Golden Horshoe Area, Ontario, Canada; R 
classoption:
  - Royal
  - times
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
---