### 2023-11-09 data-raw/00-TTS16-and-Household-Statistics.Rmd
Adding preprocessing files for newest table

---
title: "TTS Origin-Destination Data by Full-Time Employment"
output: 
---

In this notebook we process socio-economic and demographic information about households in their places of residence.

Household income ranges at the level of GTS2006 traffic analysis zones. This is the query:

```
Cross Tabulation Query Form - Person - 2016 v1.1

Row: 2006 GTA zone of household - gta06_hhld
Column: Income range of household - income

Filters:
No Filters
```
The data are in this table:

- `tts-2016-household-income-by-residence.txt`, retrieved on October 12, 2023

## Preliminaries

Load packages:
```{r load-packages}
library(dplyr) # A Grammar of Data Manipulation
library(readr) # Read Rectangular Text Data
library(tidyr) # Tidy Messy Data
```

## Number of household by income range by traffic analysis zones

Read the raw data table
```{r}
hh_sed_taz <- read_delim(file = "TTS16-data-inputs/Household-Income/tts-2016-household-income-by-taz.txt", 
                      delim = "\t", 
                      col_names = FALSE)
```

Find the position in the table where the information on work trips begins:
```{r}
idx <- which(hh_sed_taz$X1 == "Table: ")
```

Slice the table to obtain the work trips (remove heading information):
```{r}
hh_sed_taz <- hh_sed_taz |> 
  slice((idx+1):n())
```

Separate the columns and convert number of households (initially read as a character) to numeric:
```{r}
hh_sed_taz <- hh_sed_taz |>
  separate(X1, into = c("GTA06",
                        "$0 to $14999",
                        "$15000 to $39999",
                        "$40000 to $59999",
                        "$60000 to $99999",
                        "$100000 to $124999",
                        "$125000 and above",
                        "Decline / do not know"),
           sep = ",") |>
  slice_tail(n = nrow(hh_sed_taz) - 1) |>
  mutate(`$0 to $14999` = as.numeric(`$0 to $14999`),
         `$15000 to $39999` = as.numeric(`$15000 to $39999`),
         `$40000 to $59999` = as.numeric(`$40000 to $59999`),
         `$60000 to $99999` = as.numeric(`$60000 to $99999`),
         `$100000 to $124999` = as.numeric(`$100000 to $124999`),
         `$125000 and above` = as.numeric(`$125000 and above`),
         `Decline / do not know` = as.numeric(`Decline / do not know`))
```

Save to data folder:
```{r}
usethis::use_data(hh_sed_taz)
```