### 2023-11-09 data-raw/03-TTS16-GGH4-Bound-Data-Preparation-OD.Rmd

---
title: "R Notebook"
output: html_notebook
---

This notebook is to read a shapefile. The file is a zoning system used by Metrolinx and possibly compatible with TTS. I got this file from Matt Pittlock on November 2, 2023. The file is `OriginalV5_10202023`.

Load packages:
```{r}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(sf) # Simple Features for R
```

Read file:
```{r}
ggh_taz_v4 <- st_read("TTS16-data-inputs/Boundaries/OriginalV5_10202023.shp")
```

```{r}
ggh_taz_v4 <- st_make_valid(ggh_taz_v4)
```

Plot:
```{r}
ggplot() +
  geom_sf(data = st_make_valid(ggh_taz_v4))
```

Summary of object:
```{r}
summary(ggh_taz_v4)
```

Keep only useful columns:
```{r}
ggh_taz_v4 <- ggh_taz_v4 |>
  transmute(GGHV4 = taz, 
            CSDUID = csduid, 
            CSDNAME = csdname, 
            CMAUID = cmauid, 
            CMANAME = cmaname, 
            CSDUID = csduid, 
            PD = pd)
```

Transform the object so that the projection is consistent with other geographical objects.

Read a simple features object:
```{r}
load(glue::glue(here::here(), "/data/ggh_taz.rda"))
```

The object is projected as NAD83 / UTM zone 17N (epsg: 26917).

```{r}
ggh_taz_v4 <- ggh_taz_v4 |>
  st_transform(crs = st_crs(ggh_taz))
```

Plot:
```{r}
ggplot() +
  geom_sf(data = st_make_valid(ggh_taz_v4))
```

Save as simple features:
```{r}
usethis::use_data(ggh_taz_v4, 
                  overwrite = TRUE)
```